Trump vs. Harris: A 2024 US Presidential Elections Modelling Forecast

### 2024-11-04 paper/paper.qmd
added figure to discussion

For instance, a recent incident involving derogatory comments about Puerto Rico made at a Trump rally has quickly influenced the views of Puerto Rican voters in Pennsylvania. Puerto Rican voters, who make up nearly half a million people in the state, have voiced anger and disapproval, with some even reversing their voting plans [@Hill2024]. WhatsApp groups, morning radio shows, and community discussions have amplified these sentiments, illustrating how quickly backlash can spread through influential networks, particularly in swing states [@Hill2024]. Our model predicts Trump will win Pennsylvania by a margin of 0.8%, which is less than the 3.7% of Pennsylvania's Population that Puerto Ricans make up [@USCensus2020]. Situations like these could easily cause Pennsylvania's 19 electoral votes to swing the other way.

>>>>

For instance, a recent incident involving derogatory comments about Puerto Rico made at a Trump rally has quickly influenced the views of Puerto Rican voters in Pennsylvania. Puerto Rican voters, who make up nearly half a million people in the state, have voiced anger and disapproval, with some even reversing their voting plans [@Hill2024]. WhatsApp groups, morning radio shows, and community discussions have amplified these sentiments, illustrating how quickly backlash can spread through influential networks, particularly in swing states [@Hill2024]. Our model predicts Trump will win Pennsylvania by a margin of 0.8%, which is less than the 3.7% of Pennsylvania's Population that Puerto Ricans make up [@USCensus2020]. @fig-swingstates shows specifically the battleground states that are under contention. Situations like these could easily cause Pennsylvania's 19 electoral votes to swing the other way, turning from red to blue.

```{r}
#| label: fig-swingstates
#| fig-cap: "Mapped Model-predicted outcomes for 2024 elections by Swing State"
#| echo: false
#| eval: true
#| warning: false
#| message: false

# filter for swing states
battleground_states <- predictions_with_votes %>%
  select(State, TrumpPredicted, HarrisPredicted) %>%
  mutate(
    Margin = abs(TrumpPredicted - HarrisPredicted),
    Winner = case_when(
      TrumpPredicted > HarrisPredicted ~ "Donald Trump",
      HarrisPredicted > TrumpPredicted ~ "Kamala Harris",
      TRUE ~ "Tie"
    )
  ) %>%
  filter(Margin <= 4) %>%
  select(State, Winner)

battleground_states <- battleground_states %>%
  mutate(State = tolower(State))

merged_data <- states_map %>%
  left_join(battleground_states, by = c("region" = "State"))

# map that displays electoral college winners for swing states
map <- ggplot(merged_data, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = Winner), color = "white") +
  scale_fill_manual(values = c("Donald Trump" = "red", "Kamala Harris" = "blue", "NA" = "grey")) +
  labs(
    title = "2024 US Presidential Election Battleground States",
    fill = "Predicted Winner"
  ) +
  theme_void() +
  theme(legend.position = "right")
